---
title: 'Normal model'
author: 'Javier Esteban Aragoneses, Mauricio Marcos Fajgenbaun, Danyu Zhang, Daniel Alonso'
date: 'March 4th, 2021'
output: 'pdf_document'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = '#>',
fig.path = './figures/'
)
knitr::knit_engines$set(julia = JuliaCall::eng_juliacall)
options(JULIA_HOME = '/home/dreth/julia/bin')
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
library(mlbench)
```

# Introduction

This Dataset of Ozone is a dataset which contains 13 meteorological variables of , and in this work we will do i dont know, maybe a normal model. The  conclusions that we can get is that, if follows a normal  distribution and the a priori  distribution is i dont know, maybe gamma but i dont know. The distribution a posteriori is gamma? Perhaps 

```{r, echo=TRUE, warning=FALSE, message=FALSE}
data(Ozone)
```

The dataset is originally obtained from: Leo  Breiman,  Department  of  Statistics,  UC  Berkeley.   Data  used  in  Leo  Breiman  and  JeromeH. Friedman (1985), Estimating optimal transformations for multiple regression and correlation,JASA, 80, pp. 580-598.

## Variable description






# Codes 


# conclusions




